---
title: "Untitled"
author: "Christine Brown"
date: "December 1, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# This is not complete

```{r}
library(flexdashboard)
library(dplyr)
library(maps)
library(geosphere)
library( dplyr )
library( tidyr )
library( lubridate)
library( shiny )
library( ggmap )

dat <- readRDS(

        gzcon(

          url(

          "https://cdn.rawgit.com/fjsantam/Maxwell-DDM-Final-Project/master/NYC%20Bike%20Share%20Data.rds"

            )

          )

        )

# gzcon is a built-in decompressor that must be run when loading a .RDS straight from a website.


dat <- readRDS( gzcon( url( "https://cdn.rawgit.com/fjsantam/Maxwell-DDM-Final-Project/master/NYC%20Bike%20Share%20Data.rds" ) ) )
#separate date-times into dates and times
dat <- dat %>% separate( starttime, c("start.date", "start.time"), sep = " " )
dat <- dat %>% separate( stoptime, c("stop.date", "stop.time"), sep = " " )

#convert date to date format
dat$start.date <- as.Date(dat$start.date, format = "%m/%d/%Y")

#get day of week from date
dat$weekday <- weekdays(dat$start.date)
```

bbox <- make_bbox(lon=dat$start.station.longitude, lat=dat$start.station.latitude, f = .4 )

NYC <- get_map(location = bbox , maptype = "terrain-lines" )

ggmap(NYC) + geom_point(data=dat, aes(x= unique(start.station.longitude), y = unique(start.station.latitude)))






plot.new()
map( "county", region=c("new york,new york", "new york,bronx", "new york,kings", "new york,richmond") , fill=T, col="#37A4E9", bg="grey15" )
points(unique(dat$start.station.longitude), unique(dat$start.station.latitude), pch=3, cex=0.1, col="chocolate1")
