---
title: "CitiBike: When, Where, and Who?"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    social: menu
    source_code: embed #Comment this line out if run issues on SU machines
runtime: shiny
---

```{r setup, include=FALSE}
#Install Packages
library( dplyr )
library( DT )
library( flexdashboard )
library( gtable )
library( ggplot2 )
library( geosphere )
library( leaflet )
library( lubridate)
library( maps )
library( networkD3 )
library(igraph)
library( RColorBrewer )
library( scales )
library( shiny )
library( tidyr )



#Load Data
dat <- readRDS( gzcon( url( "https://cdn.rawgit.com/fjsantam/Maxwell-DDM-Final-Project/master/NYC%20Bike%20Share%20Data.rds" ) ) )

#Separate Date-Times into Dates and Times
dat <- dat %>% separate( starttime, c("start.date", "start.time"), sep = " " )
dat <- dat %>% separate( stoptime, c("stop.date", "stop.time"), sep = " " )

#Convert Date to Date Format
dat$start.date <- as.Date(dat$start.date, format = "%m/%d/%Y")

#Get Day of Week from Date
dat$weekday <- weekdays(dat$start.date, abbreviate = T)
```


Descriptive Statistics
=======================================================================

Column {data-width=650}
-----------------------------------------------------------------------

###
```{r}
source <- dat$start.station.id
target <- dat$end.station.id
name_sta <- dat$start.station.name
name_end <- dat$end.station.name
net <- data.frame(source,target, name_sta, name_end)
net <- net %>% 
  group_by(source, target, name_sta, name_end) %>% 
  summarize(count = n())
net100_links <- net %>% filter(count >=100)
net100_nodes <- select(net100_links, name_sta, source)
net100_nodes <- net100_nodes[-1]
countt <- net100_links %>% 
  group_by(name_sta)  %>% 
  summarize(avg = mean(count))
net100_links <- select(net100_links,source, target, name_sta, name_end, count)
colnames(net100_nodes)[1] <- "names"
net100_nodes <- select(net100_nodes,names)
net100_nodes <- distinct(net100_nodes)
net100_nodes$ID <- c(0:37)
net100_nodes <- subset(net100_nodes, select=c(2,1))
net100_links <- merge(net100_links, net100_nodes, by.x="name_sta", by.y="names", all.x=TRUE)
net100_links <- select(net100_links, target, name_end, ID, count)
colnames(net100_links)[3] <- "source"
net100_links <- merge(net100_links, net100_nodes, by.x="name_end", by.y="names", all.x=TRUE)
net100_links <- select(net100_links, source, ID, count)
colnames(net100_links)[2] <- "target"
net100_nodes <- mutate(net100_nodes, group = 4, size= (countt$avg-99))
net100_links$value <- (net100_links$count/10)*1.5
vertices <- net100_nodes
net100_links$source.index = match(net100_links$source, vertices$ID)-1
net100_links$target.index = match(net100_links$target, vertices$ID)-1
net100_links <- na.omit(net100_links)
tags$b(tags$h4("Most popular bike routes in NY (+100 trips a month)"))
tags$h6("Note: Node size is based on the average number trips by start station and link width is based on the number of trips between stations")
forceNetwork(Links = net100_links, Nodes = vertices,
                  Source = "source.index", Target = "target.index",
                  NodeID = "names", Nodesize = "size",
                  radiusCalculation = JS("Math.sqrt(d.nodesize)*2.5"),
                  Group = "group",
                  Value = "value",
                  charge = -100,
                  height = 600,
                  width = 800,
                  linkDistance = 40,
                  zoom = T,
                  opacity = .9,
                  fontSize=12,
                  linkColour = "#ED1330")
```

Column {data-width=400}
-----------------------------------------------------------------------

###

```{r}
sidebarLayout(
sidebarPanel(
checkboxGroupInput("checkGender", label =h5("Gender"), 
    choices = list("Male" = 1, "Female" = 2), selected = NULL),
checkboxGroupInput("checkAge", label = h5("Age Group"),
    choices = list("Post-Millennial (0-18)" =1, "Millenial (19-35)" = 2, "Gen X (36-51)" = 3, "Boomer (52-70)" = 4, "Silent (71+)" = 5), selected = NULL)
    ),
mainPanel(
renderPlot({
targetg <- input$checkGender
targeta <- input$checkAge
dat$age <- 2016-dat$birth.year
dat <- mutate(dat,agegroup=NULL)
dat$agegroup[dat$age < 19] <- 1
dat$agegroup[dat$age > 18 & dat$age < 36] <- 2
dat$agegroup[dat$age > 35 & dat$age < 52] <- 3
dat$agegroup[dat$age > 51 & dat$age < 71] <- 4
dat$agegroup[dat$age > 70] <- 5
if(is.null(targetg)) {targetg=c(0,1,2)}
if(is.null(targeta)) {targeta=c(1,2,3,4,5)}
trips <- dat %>%
  filter(gender %in% targetg) %>%
  filter(agegroup %in% targeta) %>%
  group_by( weekday ) %>%
  summarize(ntrip = n())
#order days of week
trips <- trips[c(2,6,7,5,1,3,4),]
#plot barplot (ridership by day of week)
par(oma = c(0,0,0,0),mar = c(0,0,0,0), lty = 0)
barplot <- barplot(height = trips$ntrip, 
                   xlab = "", ylim=c(0, 65000), 
                   col = "#37A4E9", 
                   axes = F)
text(x = barplot, 
     y = 0, 
     labels = trips$weekday, 
     pos = 3, 
     cex = 0.85, 
     font = 2,
     col = "white")
text(x = barplot, 
     y = trips$ntrip, 
     labels = paste(round(trips$ntrip/1000),"K", sep = ""),
     pos = 3, 
     cex = 1, 
     font = 2, 
     col = "gray0")
 }, height = 275, width = 300)
))
```


###


```{r}
test <- dat
# Converts time into values, with max value for a given hundred set at 59 and moving to the next sequential hundred digit at 60
# Example 1259, next value is 1300 NOT 1260
test$start.time <- gsub( ":", "", test$start.time )
test$start.time <- as.integer( test$start.time )
test$stop.time <- gsub( ":", "", test$stop.time )
test$stop.time <- as.integer( test$stop.time )

test$period <- cut( test$start.time, c( 0, 400, 800, 1200, 1600, 2000, 2400), labels = c("Midnight", "Early Morning", "Late Morning", "Afternoon", "Evening", "Night"), right=F)
# Saves the cuts in column "period" within the vector

# Reformat the start.date values to the "date" class
test$start.date <- as.Date(test$start.date, format = "%m/%d/%Y")
# Create a new column that has the matching Weekday name per date
test$weekday <- weekdays(test$start.date)

# Convert the Weekday names to shorthand
test$weekday <- gsub("Monday", "Mon", test$weekday, ignore.case = FALSE, fixed = FALSE)
test$weekday <- gsub("Tuesday", "Tue", test$weekday, ignore.case = FALSE, fixed = FALSE)
test$weekday <- gsub("Wednesday", "Wed", test$weekday, ignore.case = FALSE, fixed = FALSE)
test$weekday <- gsub("Thursday", "Thu", test$weekday, ignore.case = FALSE, fixed = FALSE)
test$weekday <- gsub("Friday", "Fri", test$weekday, ignore.case = FALSE, fixed = FALSE)
test$weekday <- gsub("Saturday", "Sat", test$weekday, ignore.case = FALSE, fixed = FALSE)
test$weekday <- gsub("Sunday", "Sun", test$weekday, ignore.case = FALSE, fixed = FALSE)
days.week <- c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun")
list.periods <- c( "Night", "Evening", "Afternoon", "Late Morning", "Early Morning", "Midnight")

# Create a dataframe that contains count of observations for each combination of weekdays and periods
sub2 <- test %>% group_by(weekday, period) %>% summarize( count = n() )
# Add the same station name to each row, which is effectively a summary row for that weekday-period combo
sub2$start.station.name <- " All Stations"
# Re-order the dataframe to match the parent dataframe's column order
sub2 <- sub2[,c("weekday","period","start.station.name","count")]
# Create a dataframe that contains count of observations for each combination of weekday, period, and start station
sub3 <- test %>% group_by(weekday, period, start.station.name) %>% summarize(count = n() )
# Combine the station-specific and summary dataframes into one
sub3 <- rbind(sub2, sub3)
sub3$weekday <- factor( sub3$weekday, levels = days.week )
sub3$period <- factor( sub3$period, levels = list.periods )
# Create a list of unique station names for the Shiny selection list
stations <- unique( sub3$start.station.name )
stations <- sort( as.character(stations) )
cool.colors <- c( "#27336E", "#ED1330" )
selectInput( 
           inputId='station', 
           label='Choose a Station', 
           choices = stations,
           selected = " All Stations"
         )

renderPlot({
  filtered <- filter( .data = sub3, start.station.name == input$station  )

  if( filtered$start.station.name == " All Stations"){
        ggplot( data = filtered, 
                aes( x = filtered$weekday, y = filtered$period, fill = count )
                ) + 
        geom_tile() + 
        guides( fill = F ) + 
        scale_fill_gradientn( colors=cool.colors ) +
        labs( x = "", y = "") + 
        theme( panel.border = element_blank(), 
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_rect( fill="black" )
        )  
  } else {
        ggplot( data = filtered, 
                aes( x = filtered$weekday, y = filtered$period, fill = count )
                ) + 
        geom_tile() + 
        guides( fill = F ) + 
        scale_fill_gradientn( colors=cool.colors, limits=c(0,378) ) +
        labs( x = "", y = "") + 
        theme( panel.border = element_blank(), 
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_rect( fill="black" )
        )
  }
}, height = 275)
```

Station Map
=======================================================================

```{r}
stations <- data.frame(station.name=dat$start.station.name,
                      latitude=dat$start.station.latitude,
                      longitude=dat$start.station.longitude)
stations2 <- data.frame(station.name=dat$end.station.name,
                       latitude=dat$end.station.latitude,
                       longitude=dat$end.station.longitude)
stations <- rbind(stations, stations2)
stations <- unique(stations)
dayofweek <- c("All Days", "Mon", "Tue","Wed","Thu", "Fri","Sat","Sun")
```


```{r}

  selectInput("SelectWeekday", label = h5("Choose a Weekday"), 
     choices = dayofweek, selected = "All Days")

renderLeaflet({
    if(input$SelectWeekday=="All Days") { 
      start.trips <- dat %>% 
      group_by(start.station.name) %>%
      summarize(count=n())
      
      stations <- merge(stations, start.trips, by.x="station.name", by.y="start.station.name")
      palette <- colorNumeric( palette=c("#27336E","#ED1330"), domain=stations$count)
  
      map <- leaflet(data=stations) %>%
      setView(lat=40.73029, lng=-73.99076, zoom=13) %>%
      addProviderTiles("CartoDB.Positron") %>%
      addCircles(lng=~longitude, lat=~latitude, 
             radius=~count/50, color=~palette(count), 
             popup=~station.name)
      map 
    } else {
    start.trips <- dat %>% 
    group_by(start.station.name, weekday) %>%
    filter(weekday == input$SelectWeekday) %>% 
    summarize(count=n())
    
    stations <- merge(stations, start.trips, by.x="station.name", by.y="start.station.name")
    palette <- colorNumeric( palette=c("#27336E","#ED1330"), domain=stations$count)
    map <- leaflet(data=stations) %>%
    setView(lat=40.73029, lng=-73.99076, zoom=13) %>%
    addProviderTiles("CartoDB.Positron") %>%
    addCircles(lng=~longitude, lat=~latitude, 
             radius=~count/5, color=~palette(count), 
             popup=~station.name)
    map
  }
  
  })

```

Data Download
=======================================================================

```{r}
subdat <- dat[1:200,]  
DT::datatable(subdat,
                 options = list(
                  pageLength = 25,
                  initComplete = I("function(settings, json) {alert('Done.');}")
                )
             )
```

About
=======================================================================

![](https://d21xlh2maitm24.cloudfront.net/nyc/CitiBike_Logo_p.svg?mtime=20160427183115)

This dashboard was compiled by students from the Fall 2016 Data Driven Management course at Syracuse University.

Group Members:

  - Alejandro Alfaro Aco
  - Christine Brown
  - Caitlin Eger
  - Shahzeb Khan
  - Mengyu Liu
  - Francisco Santamarina

We are grateful to CitiBike for making its data publicly available.
